---
title: "npj Precision Oncology"
author: Shengqin Su, ssu42@stanford.edu
output: html_notebook
---


### Load required library
```{r}
library(readxl)
library(survival)
library(tidyverse)
library(dplyr)
library(survminer)
library(ggplot2)
```
## Initial submission
### Survival analysis of PLK1 and Numb
```{r}
GCdata = read_excel("TCGAdata/12-2022_PLK1NUMB_v2.xlsx")%>% # CDH1 is E-cad, CDH2 is N-cad
  mutate(NumbNotch = paste(NUMB_Sta, NOTCH1_Sta, sep = ""), # LowHigh vs HighLow
         NumbCDH1 = paste(NUMB_Sta, CDH1_Sta, sep = ""), # HighHigh vs LowLow
         NumbCDH2 = paste(NUMB_Sta, CDH2_Sta, sep = ""), # LowHigh vs HighLow
         NotchCDH1 = paste(NOTCH1_Sta, CDH1_Sta, sep = ""), # LowHigh vs HighLow
         NotchCDH2 = paste(NOTCH1_Sta, CDH2_Sta, sep = ""), # HighHigh vs LowLow
         PLK1Numb = paste(PLK1_Sta, NUMB_Sta, sep = ""), # LowHigh vs HighLow
         PLK1CDH1 = paste(PLK1_Sta, CDH1_Sta, sep = ""), # HighLow vs LowHigh
         PLK1CDH2 = paste(PLK1_Sta, CDH2_Sta, sep = ""), # HighHigh vs LowLow
         ) 

GC_OSdata = GCdata %>% filter(OS_MONTHS != "[Not Available]")
GC_OSdata_NumbNotch = GC_OSdata %>% filter(NumbNotch %in% c("HighLow", "LowHigh"))
GC_OSdata_NumbCDH1 = GC_OSdata %>% filter(NumbCDH1 %in% c("HighHigh", "LowLow"))
GC_OSdata_NumbCDH2 = GC_OSdata %>% filter(NumbCDH2 %in% c("LowHigh", "HighLow"))
GC_OSdata_NotchCDH1 = GC_OSdata %>% filter(NotchCDH1 %in% c("LowHigh", "HighLow"))
GC_OSdata_NotchCDH2 = GC_OSdata %>% filter(NotchCDH2 %in% c("HighHigh", "LowLow"))
GC_OSdata_PLK1Numb = GC_OSdata %>% filter(PLK1Numb %in% c("LowHigh", "HighLow"))
GC_OSdata_PLK1CDH1 = GC_OSdata %>% filter(PLK1CDH1 %in% c("HighLow", "LowHigh"))
GC_OSdata_PLK1CDH2 = GC_OSdata %>% filter(PLK1CDH2 %in% c("HighHigh", "LowLow"))


GC_DFSdata = GCdata %>% filter(DFS_MONTHS != "[Not Available]")
GC_DFSdata_NumbNotch = GC_DFSdata %>% filter(NumbNotch %in% c("HighLow", "LowHigh"))
GC_DFSdata_NumbCDH1 = GC_DFSdata %>% filter(NumbCDH1 %in% c("HighHigh", "LowLow"))
GC_DFSdata_NumbCDH2 = GC_DFSdata %>% filter(NumbCDH2 %in% c("LowHigh", "HighLow"))
GC_DFSdata_NotchCDH1 = GC_DFSdata %>% filter(NotchCDH1 %in% c("LowHigh", "HighLow"))
GC_DFSdata_NotchCDH2 = GC_DFSdata %>% filter(NotchCDH2 %in% c("HighHigh", "LowLow"))
GC_DFSdata_PLK1Numb = GC_DFSdata %>% filter(PLK1Numb %in% c("LowHigh", "HighLow"))
GC_DFSdata_PLK1CDH1 = GC_DFSdata %>% filter(PLK1CDH1 %in% c("HighLow", "LowHigh"))
GC_DFSdata_PLK1CDH2 = GC_DFSdata %>% filter(PLK1CDH2 %in% c("HighHigh", "LowLow"))

```

```{r}
# OS Analysis
Total_list = list(GC_OSdata_NumbNotch, GC_OSdata_NumbCDH1, GC_OSdata_NumbCDH2, GC_OSdata_NotchCDH1, GC_OSdata_NotchCDH2, GC_OSdata_PLK1Numb, GC_OSdata_PLK1CDH1, GC_OSdata_PLK1CDH2)
List_name = c("NumbNotch","NumbCDH1", "NumbCDH2", "NotchCDH1", "NotchCDH2", "PLK1Numb", "PLK1CDH1", "PLK1CDH2")
#names(Total_list[[1]])
for (i in 1:8) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$OS_MONTHS),event = as.numeric(temp_frame$OS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("OS_",List_name[i],".pdf", sep = ""), p$plot, dpi = 300)
}
```

```{r}
# DFS analysis
Total_list = list(GC_DFSdata_NumbNotch, GC_DFSdata_NumbCDH1, GC_DFSdata_NumbCDH2, GC_DFSdata_NotchCDH1, GC_DFSdata_NotchCDH2, GC_DFSdata_PLK1Numb, GC_DFSdata_PLK1CDH1, GC_DFSdata_PLK1CDH2)
List_name = c("NumbNotch","NumbCDH1", "NumbCDH2", "NotchCDH1", "NotchCDH2", "PLK1Numb", "PLK1CDH1", "PLK1CDH2")
#names(Total_list[[1]])
for (i in 1:7) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$DFS_MONTHS),event = as.numeric(temp_frame$DFS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("DFS_", List_name[i],".pdf", sep = ""), p$plot)
}
```

## 020723: Major revision from reviewer.
### Survival analysis
Generate this: 
1. PLK1-H/Numb-L/NOTCH-H – PLK1-L/Numb-H/NOTCH-L
2. PLK1-H/CDH1-L/Numb-L – PLK1-L/CDH1-H/Numb-H
3. PLK1-H/CDH2-H/Numb-L – PLK1-L/CDH2-L/Numb-H

```{r}
GCdata = read_excel("TCGAdata/12-2022_PLK1NUMB_v2.xlsx")%>% # CDH1 is E-cad, CDH2 is N-cad
  mutate(PLK1NumbNotch = paste(PLK1_Sta, NUMB_Sta, NOTCH1_Sta, sep = ""), # LowHighLow vs HighLowHigh
         PLK1NumbCDH1 = paste(PLK1_Sta, NUMB_Sta, CDH1_Sta, sep = ""), # LowHighHigh vs HighLowLow
         PLK1NumbCDH2 = paste(PLK1_Sta, NUMB_Sta, CDH2_Sta, sep = ""), # HighLowHigh vs LowHighLow
         ) # LowLow vs HighHigh

GC_OSdata = GCdata %>% filter(OS_MONTHS != "[Not Available]")
GC_OSdata_PLK1NumbNotch = GC_OSdata %>% filter(PLK1NumbNotch %in% c("LowHighLow", "HighLowHigh"))
GC_OSdata_PLK1NumbCDH1 = GC_OSdata %>% filter(PLK1NumbCDH1 %in% c("LowHighHigh", "HighLowLow"))
GC_OSdata_PLK1NumbCDH2 = GC_OSdata %>% filter(PLK1NumbCDH2 %in% c("HighLowHigh", "LowHighLow"))

GC_DFSdata = GCdata %>% filter(DFS_MONTHS != "[Not Available]")
GC_DFSdata_PLK1NumbNotch = GC_DFSdata %>% filter(PLK1NumbNotch %in% c("LowHighLow", "HighLowHigh"))
GC_DFSdata_PLK1NumbCDH1 = GC_DFSdata %>% filter(PLK1NumbCDH1 %in% c("LowHighHigh", "HighLowLow"))
GC_DFSdata_PLK1NumbCDH2 = GC_DFSdata %>% filter(PLK1NumbCDH2 %in% c("HighLowHigh", "LowHighLow"))



```


```{r}
# OS plot
Total_list = list(GC_OSdata_PLK1NumbNotch, GC_OSdata_PLK1NumbCDH1, GC_OSdata_PLK1NumbCDH2)
List_name = c("PLK1NumbNotch","PLK1NumbCDH1", "PLK1NumbCDH2")
#names(Total_list[[1]])
for (i in 1:3) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$OS_MONTHS),event = as.numeric(temp_frame$OS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("OS_", List_name[i],".pdf", sep = ""), p$plot)
}
```

```{r}
# DFS plot
Total_list = list(GC_DFSdata_PLK1NumbNotch, GC_DFSdata_PLK1NumbCDH1, GC_DFSdata_PLK1NumbCDH2)
List_name = c("PLK1NumbNotch","PLK1NumbCDH1", "PLK1NumbCDH2")
#names(Total_list[[1]])
for (i in 1:3) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$DFS_MONTHS),event = as.numeric(temp_frame$DFS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("DFS_", List_name[i],".pdf", sep = ""), p$plot)
}
```

### Boxplot for NOTCH expression
```{r}
GCdata = read_excel("TCGAdata/12-2022_PLK1NUMB_v2.xlsx")%>% # CDH1 is E-cad, CDH2 is N-cad
  mutate(
         PLK1Numb = paste(PLK1_Sta, NUMB_Sta, sep = ""),
         ) # LowLow vs HighHigh

GC_PLK1_H = GCdata %>% filter(PLK1_Sta %in% c("High")) %>% select(NOTCH1_Median, PLK1_Sta) %>% mutate(PLK1_Sta = str_replace(PLK1_Sta, "High", "PLK1_H")) %>% rename("Condition" = "PLK1_Sta")

GC_PLK1_L = GCdata %>% filter(PLK1_Sta %in% c("Low")) %>% select(NOTCH1_Median, PLK1_Sta) %>% mutate(PLK1_Sta = str_replace(PLK1_Sta, "Low", "PLK1_L")) %>% rename("Condition" = "PLK1_Sta")

GC_Numb_H = GCdata %>% filter(NUMB_Sta %in% c("High")) %>% select(NOTCH1_Median, NUMB_Sta) %>% mutate(NUMB_Sta = str_replace(NUMB_Sta, "High", "NUMB_H")) %>% rename("Condition" = "NUMB_Sta")

GC_Numb_L = GCdata %>% filter(NUMB_Sta %in% c("Low")) %>% select(NOTCH1_Median, NUMB_Sta) %>% mutate(NUMB_Sta = str_replace(NUMB_Sta, "Low", "NUMB_L")) %>% rename("Condition" = "NUMB_Sta")

GC_PLK1Numb_HL = GCdata %>% filter(PLK1Numb %in% c("HighLow")) %>% select(NOTCH1_Median, PLK1Numb) %>% mutate(PLK1Numb = str_replace(PLK1Numb, "HighLow", "PLK1Numb_HL")) %>% rename("Condition" = "PLK1Numb")

GC_PLK1Numb_LH = GCdata %>% filter(PLK1Numb %in% c("LowHigh")) %>% select(NOTCH1_Median, PLK1Numb) %>% mutate(PLK1Numb = str_replace(PLK1Numb, "LowHigh", "PLK1Numb_LH")) %>% rename("Condition" = "PLK1Numb")


```

Compare Notch1 expression between PLK1_NUMB
```{r}

NOTCH1_condition = rbind(GC_PLK1Numb_HL, GC_PLK1Numb_LH)
p2 = ggplot(NOTCH1_condition, aes(x = Condition, y = log(NOTCH1_Median), color = Condition)) + geom_boxplot() + stat_compare_means(method = "t.test")
ggsave(file = "Notch_exp_btwn_allgroup.pdf", plot = p2, dpi = 300)


NOTCH1_condition = rbind(GC_PLK1_H, GC_PLK1_L, GC_Numb_H, GC_Numb_L, GC_PLK1Numb_HL, GC_PLK1Numb_LH)
p2 = ggplot(NOTCH1_condition, aes(x = Condition, y = log(NOTCH1_Median), color = Condition)) + geom_boxplot() + stat_compare_means(method = "anova")
ggsave(file = "Notch_exp_btwn_allgroup.pdf", plot = p2, dpi = 300)
```



### Numb isoform analysis: Survival analysis on the expression of PLK1 and Numb1-4
```{r}
GCdata = read_excel("TCGAdata/12-2022_PLK1NUMB_v2.xlsx")%>% # CDH1 is E-cad, CDH2 is N-cad
  mutate(NUMB4Notch = paste(NUMB4_Sta, NOTCH1_Sta, sep = ""), # LowHigh vs HighLow
         NUMB4CDH1 = paste(NUMB4_Sta, CDH1_Sta, sep = ""), # HighHigh vs LowLow
         NUMB4CDH2 = paste(NUMB4_Sta, CDH2_Sta, sep = ""), # LowHigh vs HighLow
         PLK1NUMB4 = paste(PLK1_Sta, NUMB4_Sta, sep = ""), # LowHigh vs HighLow
         ) 

GC_OSdata = GCdata %>% filter(OS_MONTHS != "[Not Available]")
GC_OSdata_NUMB4Notch = GC_OSdata %>% filter(NUMB4Notch %in% c("LowHigh", "HighLow"))
GC_OSdata_NUMB4CDH1 = GC_OSdata %>% filter(NUMB4CDH1 %in% c("HighHigh", "LowLow"))
GC_OSdata_NUMB4CDH2 = GC_OSdata %>% filter(NUMB4CDH2 %in% c("LowHigh", "HighLow"))
GC_OSdata_PLK1NUMB4 = GC_OSdata %>% filter(PLK1NUMB4 %in% c("LowHigh", "HighLow"))

GC_DFSdata = GCdata %>% filter(DFS_MONTHS != "[Not Available]")
GC_DFSdata_NUMB4Notch = GC_DFSdata %>% filter(NUMB4Notch %in% c("LowHigh", "HighLow"))
GC_DFSdata_NUMB4CDH1 = GC_DFSdata %>% filter(NUMB4CDH1 %in% c("HighHigh", "LowLow"))
GC_DFSdata_NUMB4CDH2 = GC_DFSdata %>% filter(NUMB4CDH2 %in% c("LowHigh", "HighLow"))
GC_DFSdata_PLK1NUMB4 = GC_DFSdata %>% filter(PLK1NUMB4 %in% c("LowHigh", "HighLow"))


```

```{r}
# OS Plot
Total_list = list(GC_OSdata_NUMB4Notch, GC_OSdata_NUMB4CDH1, GC_OSdata_NUMB4CDH2, GC_OSdata_PLK1NUMB4)
#, GC_OSdata_NotchCDH2, GC_OSdata)
List_name = c("NUMB4Notch","NUMB4CDH1", "NUMB4CDH2", "PLK1NUMB4")
#names(Total_list[[1]])
for (i in 1:4) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$OS_MONTHS),event = as.numeric(temp_frame$OS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("OS_", List_name[i],".pdf", sep = ""), p$plot, dpi = 300)
}
```

```{r}
# DFS Plot
Total_list = list(GC_DFSdata_NUMB4Notch, GC_DFSdata_NUMB4CDH1, GC_DFSdata_NUMB4CDH2, GC_DFSdata_PLK1NUMB4)
#, GC_DFSdata_NotchCDH2, GC_DFSdata)
List_name = c("NUMB4Notch","NUMB4CDH1", "NUMB4CDH2", "PLK1NUMB4")
#names(Total_list[[1]])
for (i in 1:4) {
  temp_frame = as.data.frame(Total_list[[i]])
  temp_name = List_name[i]
  modform <- as.formula(paste("Surv(time = as.numeric(temp_frame$DFS_MONTHS),event = as.numeric(temp_frame$DFS_STATUS))", List_name[i], sep = " ~ "))
  substitute(survfit(modform, data = temp_frame), list(modform = modform))
  funny = eval(substitute(survfit(modform, data = temp_frame), list(modform = modform)))
  p <- ggsurvplot(funny, data = temp_frame, pval = TRUE)
  ggsave(file = paste("DFS_", List_name[i],".pdf", sep = ""), p$plot, dpi = 300)
}
```


### Numb isoform analysis: comparing Numb1-4 expression between TCGA vs GTEx
```{r}
Numb_iso = read_excel("TCGAdata/Numb_isoforms_TPM_TCGA.xlsx", sheet = "Sheet4")

Numb1_GTEx_vs_TCGA = ggplot(Numb_iso, aes(x = Type, y = log2(`Numb-1`+1),  color = Type)) + geom_boxplot() + 
  stat_compare_means(method = "t.test")
Numb2_GTEx_vs_TCGA = ggplot(Numb_iso, aes(x = Type, y = log2(`Numb-2`+1),  color = Type)) + geom_boxplot() + 
  stat_compare_means(method = "t.test")
Numb3_GTEx_vs_TCGA = ggplot(Numb_iso, aes(x = Type, y = log2(`Numb-3`+1),  color = Type)) + geom_boxplot() + 
  stat_compare_means(method = "t.test")
Numb4_GTEx_vs_TCGA = ggplot(Numb_iso, aes(x = Type, y = log2(`Numb-4`+1),  color = Type)) + geom_boxplot() + 
  stat_compare_means(method = "t.test")

ggsave("Numb1_GTEx_vs_TCGA.pdf", plot = Numb1_GTEx_vs_TCGA, dpi = 300)
ggsave("Numb2_GTEx_vs_TCGA.pdf", plot = Numb2_GTEx_vs_TCGA, dpi = 300)
ggsave("Numb3_GTEx_vs_TCGA.pdf", plot = Numb3_GTEx_vs_TCGA, dpi = 300)
ggsave("Numb4_GTEx_vs_TCGA.pdf", plot = Numb4_GTEx_vs_TCGA, dpi = 300)
```

