---
title: "MZB1-in-melanoma"
author: "Shengqin Su"
date: "2024-11-07"
output: html_notebook
---

### Load library
```{r}
library(dplyr)
library(ggplot2)
library(Seurat)
library(limma)
library(edgeR)
library(readxl)
library(reshape2)
library(survival)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(pathview)
```


### Data cleanning on TCGA and GTEx data
```{r}
SKSun_GTEx = read.table("data/gene_tpm_2017-06-05_v8_skin_sun_exposed_lower_leg.gct", header = TRUE, sep = "\t", check.names = F)

SKNoSun_GTEx = read.table("data/gene_tpm_2017-06-05_v8_skin_not_sun_exposed_suprapubic.gct", header = TRUE, sep = "\t", check.names = F)

SK_GTEx = merge(SKSun_GTEx, SKNoSun_GTEx, by = "Name") %>% select(-id.y, -Description.y, -id.x) %>% rename(Description = Description.x)

write.table(SK_GTEx, file = "data/2024-06-02_Skin_GTEx_Sun_noSun.txt", sep = "\t", row.names = F, quote = F)

SKCM_TCGA = read.table("data/TCGA-SKCM.htseq_fpkm.tsv", row.names = 1, header = TRUE, sep = "\t", check.names = F)

SKCM_TCGA_FPKM = 2^SKCM_TCGA - 1

SKCM_TCGA_TPM = SKCM_TCGA_FPKM
# Convert FPKM to TPM
for (col in colnames(SKCM_TCGA_TPM)) {
  sum_fpkm <- sum(SKCM_TCGA_TPM[[col]])  # Use double brackets to get the column correctly
  SKCM_TCGA_TPM[[col]] <- (SKCM_TCGA_TPM[[col]] / sum_fpkm) * 1e6
}

write.table(cbind(Gene = row.names(SKCM_TCGA_TPM), SKCM_TCGA_TPM), file = "data/2024-06-02_SKCM_TPM.txt", sep = "\t", row.names = F, quote = F)

```

### Perform DESeq between primary vs metastatic in TCGA
```{r}
#Load the data and prepare the data
SKCM_TCGA_TPM = read.table("data/2024-06-02_SKCM_TPM.txt", row.names = 1, check.names = F, sep = "\t", header = T)

primary_cols <- grep("01[A-Z]$", colnames(SKCM_TCGA_TPM), value = TRUE)
metastasis_cols <- grep("06[A-Z]$", colnames(SKCM_TCGA_TPM), value = TRUE)


SKCM_Pr_TPM = SKCM_TCGA_TPM[, primary_cols]
SKCM_Mt_TPM = SKCM_TCGA_TPM[, metastasis_cols]

combined_tpm_data <- cbind(SKCM_Pr_TPM, SKCM_Mt_TPM)
# Filter the data

average_threshold <- 0.5
zero_threshold <- 0.8  # e.g., genes must be expressed in at least 20% of samples

# Calculate the average expression of each gene across all samples
average_expression <- rowMeans(combined_tpm_data)

# Calculate the proportion of samples with zero expression for each gene
proportion_zero <- rowMeans(combined_tpm_data == 0)

# Filter out genes with average expression below the threshold and those with high proportion of zeroes
filtered_tpm_data <- combined_tpm_data[average_expression > average_threshold & proportion_zero < zero_threshold, ]

# Perform DEG
group <- factor(c(rep("Pt", ncol(SKCM_Pr_TPM)), rep("Mt", ncol(SKCM_Mt_TPM))))
design <- model.matrix(~0 + group)
colnames(design) <- levels(group)
contrast_matrix <- makeContrasts(Mt_vs_Pt = Mt - Pt, levels = design)


fit <- lmFit(filtered_tpm_data, design)
fit2 <- contrasts.fit(fit, contrast_matrix)

fit2 <- eBayes(fit2)

results <- topTable(fit2, coef = "Mt_vs_Pt", adjust.method = "BH", number = Inf)
write.csv(results, "data/2024-06-04_MT_vs_PT_nolog_DEG_contrast.csv")

```

## Perform DESeq between normal GTEx vs primary TCGA
```{r}
GTEx = read.table("data/2024-06-02_Skin_GTEx_Sun_noSun.txt", row.names = 1, check.names = F, sep = "\t", header = T)

ENSG_Genename = cbind(rownames(GTEx), GTEx[,1])
colnames(ENSG_Genename) = c("ENSG", "Symbol")
write.table(ENSG_Genename, file = "data/2024-06-20_ENSG_to_Symbol.txt", sep = "\t", row.names = F, quote = F)

GTEx_rename = cbind(ENSG = sapply(strsplit(rownames(GTEx), "\\."), `[`, 1), GTEx)
SKCM_Pr_TPM_rename = cbind(ENSG = sapply(strsplit(rownames(SKCM_Pr_TPM), "\\."), `[`, 1), SKCM_Pr_TPM)
GTEx_TCGA_common = intersect(GTEx_rename$ENSG, SKCM_Pr_TPM_rename$ENSG)

GTEx_rename_common = GTEx_rename %>% filter(ENSG %in% GTEx_TCGA_common)
SKCM_Pr_rename_common = SKCM_Pr_TPM_rename %>% filter(ENSG %in% GTEx_TCGA_common)

GTEx_rename_common_sum = as.data.table(GTEx_rename_common)[, lapply(.SD, sum, na.rm = TRUE), by = .(ENSG, Description)]

GTEx_TCGA_Pr = merge(SKCM_Pr_rename_common, GTEx_rename_common_sum, by = "ENSG" ) %>%
  select(-Description)

rownames(GTEx_TCGA_Pr) = GTEx_TCGA_Pr$ENSG
GTEx_TCGA_Pr = GTEx_TCGA_Pr %>% select(-ENSG)

average_threshold <- 0.5
zero_threshold <- 0.8  # e.g., genes must be expressed in at least 10% of samples

# Calculate the average expression of each gene across all samples
average_expression <- rowMeans(GTEx_TCGA_Pr)

# Calculate the proportion of samples with zero expression for each gene
proportion_zero <- rowMeans(GTEx_TCGA_Pr == 0)

# Filter out genes with average expression below the threshold and those with high proportion of zeroes
GTEx_TCGA_filtered <- GTEx_TCGA_Pr[average_expression > average_threshold & proportion_zero < zero_threshold, ]

GTEx_filtered = GTEx_TCGA_filtered %>% select(starts_with("GTEx"))
TCGA_filtered = GTEx_TCGA_filtered %>% select(starts_with("TCGA"))

#write.table(cbind(ENSG = rownames(GTEx_filtered), GTEx_filtered), file = "data/2024-06-09_GTEx_filtered.txt", sep = "\t", row.names = F, quote = F)

#write.table(cbind(ENSG = rownames(TCGA_filtered), TCGA_filtered), file = "data/2024-06-09_SKCM_Pr_filtered.txt", sep = "\t", row.names = F, quote = F)

GTEx_filtered_t = GTEx_filtered %>% t() %>% as.data.frame()

GTEx_filtered_t = cbind(Patient_ID = sapply(strsplit(rownames(GTEx_filtered_t), "-"), `[`, 2), GTEx_filtered_t)

GTEx_ave = GTEx_filtered_t %>% group_by(Patient_ID) %>%
    summarize(across(everything(), mean)) %>%
  ungroup()

GTEx_ave = as.data.frame(GTEx_ave)
GTEx_ave = GTEx_ave[,-Patient_ID]
row.names(GTEx_ave) = paste0("GTEx-", GTEx_ave$Patient_ID)

GTEx_ave = GTEx_ave %>% select(-Patient_ID)

GTEx_ave_t = GTEx_ave %>% t() %>% as.data.frame()

GTEx_TCGA_final = cbind(GTEx_ave_t, TCGA_filtered)
backup  = GTEx_TCGA_final

# Normalize to TPM
for (col in colnames(GTEx_TCGA_final)) {
  sum_fpkm <- sum(GTEx_TCGA_final[[col]])  # Use double brackets to get the column correctly
  GTEx_TCGA_final[[col]] <- (GTEx_TCGA_final[[col]] / sum_fpkm) * 1e6
}

write.table(cbind(ENSG = rownames(GTEx_TCGA_final), GTEx_TCGA_final), file = "data/2024-06-09_GTEx_SKCM_Pr_TPM.txt", sep = "\t", row.names = F, quote = F)

```


### Get common DEGs
```{r}
Pt_vs_normal = read.table("data/2024-06-09_PT_vs_GTEx_nolog_DEG_contrast.csv", sep = ",", header = T)

PT_DEG = Pt_vs_normal %>% filter((logFC >= 1 | logFC <= -1) & adj.P.Val < 0.05)

PT_DEG_ENSG = PT_DEG$X
Mt_vs_Pt = read.table("data/2024-06-04_MT_vs_PT_nolog_DEG_contrast.csv", sep = ",", header = T)

MT_DEG = Mt_vs_Pt %>% filter((logFC >= 1 | logFC <= -1) & adj.P.Val < 0.05)

MT_DEG_ENSG = sub("\\..*", "", MT_DEG$X)

common_DEG <- intersect(PT_DEG_ENSG, MT_DEG_ENSG)
length(common_DEG)

```

### TCGA Survival analysis
```{r}
SKCM_Exp = read.table("data/2024-06-02_SKCM_TPM.txt", sep = "\t", header = T, row.names = 1)

average_threshold <- 0.5
zero_threshold <- 0.8  # e.g., genes must be expressed in at least 10% of samples

# Calculate the average expression of each gene across all samples
average_expression <- rowMeans(SKCM_Exp)

# Calculate the proportion of samples with zero expression for each gene
proportion_zero <- rowMeans(SKCM_Exp == 0)

# Filter out genes with average expression below the threshold and those with high proportion of zeroes
SKCM_Exp_filtered <- SKCM_Exp[average_expression > average_threshold & proportion_zero < zero_threshold, ]

# Convert FPKM to TPM
for (col in colnames(SKCM_Exp_filtered)) {
  sum_col <- sum(SKCM_Exp_filtered[[col]])  # Use double brackets to get the column correctly
  SKCM_Exp_filtered[[col]] <- (SKCM_Exp_filtered[[col]] / sum_col) * 1e6
}

SKCM_Exp_T = SKCM_Exp_filtered %>% (t) %>% as.data.frame()

SKCM_Exp_T = cbind(PatientID = substr(row.names(SKCM_Exp_T), 1, nchar(row.names(SKCM_Exp_T)) - 4), SKCM_Exp_T) %>% group_by(PatientID) %>%
  summarize(across(everything(), mean)) %>%
  ungroup() %>% as.data.frame()


SKCM_clinial = read.table("data/skcm_tcga_pan_can_atlas_2018/data_clinical_patient.txt", skip = 4, sep = "\t")

colnames(SKCM_clinial) = SKCM_clinial[1,]
SKCM_clinial = SKCM_clinial[-1,]
SKCM_clinial$PATIENT_ID = gsub("-", ".", SKCM_clinial$PATIENT_ID)
SKCM_clinial_OS = SKCM_clinial %>% dplyr::select(PATIENT_ID,OS_STATUS, OS_MONTHS) %>% mutate(OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0)) %>% filter(OS_MONTHS > 0)


SKCM_Exp_OS = merge(SKCM_Exp_T, SKCM_clinial_OS, by.x = "PatientID", by.y = "PATIENT_ID", all = F)



# Assuming your data frame is named df
# Remove the PatientID column as it's not needed for the analysis
data <- SKCM_Exp_OS %>% dplyr::select(-PatientID)

# Extract the OS_Status and OS_Months columns
OS_Status <- data$OS_STATUS
OS_Months <- as.numeric(data$OS_MONTHS)

# Create an empty data frame to store the results
results <- data.frame(Gene = character(),
                      coef = numeric(),
                      exp_coef = numeric(),
                      se = numeric(),
                      z = numeric(),
                      p = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each gene column and perform Cox regression
for (gene in colnames(data)[-c((ncol(data)-1), ncol(data))]) {
  # Create a Cox model formula
  formula <- as.formula(paste("Surv(OS_Months, OS_Status) ~", gene))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = data)
  
  # Extract the summary of the Cox model
  summary_cox <- summary(cox_model)
  
  # Extract the coefficients and other statistics
  coef <- summary_cox$coefficients[1, "coef"]
  exp_coef <- summary_cox$coefficients[1, "exp(coef)"]
  se <- summary_cox$coefficients[1, "se(coef)"]
  z <- summary_cox$coefficients[1, "z"]
  p <- summary_cox$coefficients[1, "Pr(>|z|)"]
  
  # Append the results to the results data frame
  results <- rbind(results, data.frame(Gene = gene, coef = coef, exp_coef = exp_coef, se = se, z = z, p = p))
}

# Print the summarized results
print(results)

results$adj_p <- p.adjust(results$p, method = "BH")

write.table(results , "data/2024-06-21_SKCM_OS_Cox.txt", row.names = F, quote = F, sep = "\t")

```

### Make survival plot
```{r}
GoI_ENSG = c("ENSG00000129455", "ENSG00000155307", "ENSG00000170476", "ENSG00000123338", "ENSG00000105246")

TCGA_OS = read.table("data/2024-06-21_SKCM_OS_Cox.txt", sep = "\t", header = T) 
TCGA_OS$Gene <- sub("\\..*", "", TCGA_OS$Gene)

TCGA_OS = TCGA_OS %>% filter(Gene %in% GoI_ENSG)

# Calculate -log10(adj_p)
TCGA_OS <- TCGA_OS %>%
  mutate(HR = exp_coef,
         lower_CI = exp(coef - 1.96 * se),
         upper_CI = exp(coef + 1.96 * se),
         neg_log10_adj_p = -log10(adj_p)) %>% left_join(Symbol_ENSG, by = c("Gene" = "ENSG"))
# Specify the order of genes
gene_order <- c("SAMSN1", "NCKAP1L", "MZB1", "KLK8", "EBI3")

# Convert Gene column to a factor with the specified order
TCGA_OS$GeneSymbol <- factor(TCGA_OS$GeneSymbol, levels = gene_order)

# HR plot
HR_plot =ggplot(TCGA_OS, aes(x = GeneSymbol, y = HR, color = neg_log10_adj_p)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), width = 0.2) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Hazard Ratio (HR) Plot",
       y = "Hazard Ratio (HR)",
       color = "-log10(adj_p)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


### Print Venn diagram
```{r}
OS = read.table("data/2024-06-21_SKCM_OS_Cox.txt", sep = "\t", header = T) %>% filter(adj_p < 0.05)

PT_DEG = read.table("data/2024-06-09_PT_vs_GTEx_nolog_DEG_contrast.csv", sep = ",", header = T) %>% filter(adj.P.Val < 0.05, abs(logFC) > 1)

MT_DEG = read.table("data/2024-06-04_MT_vs_PT_nolog_DEG_contrast.csv", sep = ",", header = T) %>% filter(adj.P.Val < 0.05, abs(logFC) > 1) 




ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

OS_ENSG <- sapply(strsplit(OS$Gene, "\\."), `[`, 1)

PT_ENSG = PT_DEG$X
MT_ENSG = sapply(strsplit(MT_DEG$X, "\\."), `[`, 1)


DEG_genes = intersect(PT_ENSG, MT_ENSG)
# Set up the Ensembl connection




DEG_genes = intersect(PT_ENSG, MT_ENSG)

OS_DEG_genes = intersect(OS_ENSG, DEG_genes)


# Query the Ensembl database to get gene symbols
OS_DEG_gene_sybol <- getBM(attributes = c('ensembl_gene_id', 'hgnc_symbol'), 
               filters = 'ensembl_gene_id', 
               values = OS_DEG_genes, 
               mart = ensembl)

write.table(OS_DEG_gene_sybol, "data/2024-06-21_OS_DEG_gene.txt", sep = "\t", quote = F, row.names = F)

# Creating the list of sets
sets <- list(
  "Primary_vs_normal" = PT_ENSG,
  "Metastatic_vs_primary" = MT_ENSG,
  "OS" = OS_ENSG
)

# Creating the Venn diagram
Melaonoma_Venn = ggvenn(sets, 
       fill_color = c("red", "green", "blue"),
       stroke_size = 0.5, 
       set_name_size = 6)
ggsave("analysis/2024-06-21_Melanoma_Venn.pdf", Melaonoma_Venn, dpi = 300)

# Make OS vocano plot
OS = read.table("data/2024-06-21_SKCM_OS_Cox.txt", sep = "\t", header = T)
OS$neg_log_p <- -log10(OS$adj_p)

# Create the volcano plot
# Create the volcano plot with specific colors
p_cox = ggplot(OS, aes(x = exp_coef, y = neg_log_p)) +
  geom_point(aes(color = adj_p < 0.05), alpha = 0.5, size = 0.1) + 
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "blue")) +
  theme_minimal() +
  labs(title = "Volcano Plot",
       x = "Hazard Ratio (HR)",
       y = "-log10 Ajusted P-Value") +
  theme(plot.title = element_text(hjust = 0.5))

ggsave("analysis/2024-08-11_Melanoma_cox.pdf", p_cox, dpi = 300)


```


### GO analysis
```{r}
# Perform GO enrichment analysis using ENSG IDs for BP ontology
ego_bp <- enrichGO(gene          = OS_DEG_genes,
                   OrgDb         = org.Hs.eg.db,
                   keyType       = 'ENSEMBL',
                   ont           = "BP",
                   pAdjustMethod = "BH",
                   qvalueCutoff  = 0.05,
                   readable      = TRUE)

ego_mf <- enrichGO(gene          = OS_DEG_genes,
                   OrgDb         = org.Hs.eg.db,
                   keyType       = 'ENSEMBL',
                   ont           = "MF",
                   pAdjustMethod = "BH",
                   qvalueCutoff  = 0.05,
                   readable      = TRUE)

ego_cc <- enrichGO(gene          = OS_DEG_genes,
                   OrgDb         = org.Hs.eg.db,
                   keyType       = 'ENSEMBL',
                   ont           = "CC",
                   pAdjustMethod = "BH",
                   qvalueCutoff  = 0.05,
                   readable      = TRUE)

# View the results
head(ego_bp)

# Plot the dotplot for Biological Process ontology
plot_bp =  dotplot(ego_bp, showCategory = 20) + ggtitle("GO Enrichment Analysis - Biological Process")
ggsave("analysis/2024-06-21_GO_BP_top20.pdf", plot_bp)

plot_mf = dotplot(ego_mf, showCategory = 20) + ggtitle("GO Enrichment Analysis - Molecular Function")
ggsave("analysis/2024-06-21_GO_MF_top20.pdf", plot_mf)


plot_cc = dotplot(ego_cc, showCategory = 20) + ggtitle("GO Enrichment Analysis - Cellular Component")
ggsave("analysis/2024-06-21_GO_CC_top20.pdf", plot_cc)

significant_go_bp <- ego_bp@result[ego_bp@result$qvalue < 0.05, ]

```



### Print volcano plot
```{r}
fivegenes = c("ENSG00000105246",
  "ENSG00000129455",
  "ENSG00000170476",
  "ENSG00000123338",
  "ENSG00000155307")

PT_DEG = read.table("data/2024-06-09_PT_vs_GTEx_nolog_DEG_contrast.csv", sep = ",", header = T)

# Correct the mutate step to ensure capping is applied correctly
PT_DEG <- PT_DEG %>% mutate(
  significance = ifelse(adj.P.Val < 0.05 & abs(logFC) > 1, "Significant", "Not Significant"),
  capped_logFC = ifelse(logFC < -50, -50, ifelse(logFC > 50, 50, logFC)),
  capped_neg_log10_p_adj = ifelse(-log10(adj.P.Val) < 0, 0, ifelse(-log10(adj.P.Val) > 50, 50, -log10(adj.P.Val)))
)

# Create the volcano plot with the correct y-axis mapping
PT_volcano_plot <- ggplot(PT_DEG, aes(x = capped_logFC, y = capped_neg_log10_p_adj, color = significance)) +
  geom_point(alpha = 0.4, size = 0.2) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "Volcano Plot",
    x = "Log Fold Change (logFC)",
    y = "-Log10 Adjusted P-Value",
    color = "Significance"
  ) +
  xlim(-50, 50) +
  ylim(0, 50) +
  theme(legend.position = "top")

ggsave("analysis/2024-06-22_PT_vs_Normal_volcano.pdf", PT_volcano_plot)


MT_DEG = read.table("data/2024-06-04_MT_vs_PT_nolog_DEG_contrast.csv", sep = ",", header = T)

# Correct the mutate step to ensure capping is applied correctly
MT_DEG <- MT_DEG %>% mutate(
  significance = ifelse(adj.P.Val < 0.05 & abs(logFC) > 1, "Significant", "Not Significant"),
  capped_logFC = ifelse(logFC < -25, -25, ifelse(logFC > 25, 25, logFC)),
  capped_neg_log10_p_adj = ifelse(-log10(adj.P.Val) < 0, 0, ifelse(-log10(adj.P.Val) > 20, 20, -log10(adj.P.Val)))
)

# Create the volcano plot with the correct y-axis mapping
MT_volcano_plot <- ggplot(MT_DEG, aes(x = capped_logFC, y = capped_neg_log10_p_adj, color = significance)) +
  geom_point(alpha = 0.4, size = 0.2) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "Volcano Plot",
    x = "Log Fold Change (logFC)",
    y = "-Log10 Adjusted P-Value",
    color = "Significance"
  ) +
  xlim(-25, 25) +
  ylim(0, 20) +
  theme(legend.position = "top")

ggsave("analysis/2024-06-22_MT_vs_PT_volcano.pdf", MT_volcano_plot)
# Display the plot
print(PT_volcano_plot)


Symbol_ENSG <- data.frame(
  GeneSymbol = c("EBI3", "KLK8", "MZB1", "NCKAP1L", "SAMSN1"),
  ENSG = c("ENSG00000105246", "ENSG00000129455", "ENSG00000170476", "ENSG00000123338", "ENSG00000155307"),
  stringsAsFactors = FALSE
)


MT_DEG_5gene = MT_DEG %>% mutate(X = sapply(strsplit(X, "\\."), `[`, 1) ) %>% filter(X %in% fivegenes)  %>% left_join(Symbol_ENSG, by = c("X" = "ENSG"))

MT_DEG_plot = ggplot(MT_DEG_5gene, aes(x = logFC, y = GeneSymbol, color = -log10(adj.P.Val))) +
  geom_point(size = 3) +
  #scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Dot Plot of DEGs in Metastatic vs primary melanoma",
    x = "Log Fold Change (logFC)",
    y = "Gene Symbol",
    color = "-Log10 Adjusted P-Value"
  ) +
  theme(legend.position = "top")
ggsave("analysis/2024-06-23_Metastatic_vs_Primary.pdf", MT_DEG_plot)

PT_DEG_5gene = PT_DEG %>% mutate(X = sapply(strsplit(X, "\\."), `[`, 1) ) %>% filter(X %in% fivegenes)  %>% left_join(Symbol_ENSG, by = c("X" = "ENSG"))

PT_DEG_plot = ggplot(PT_DEG_5gene, aes(x = logFC, y = GeneSymbol, color = -log10(adj.P.Val))) +
  geom_point(size = 3) +
  #scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Dot Plot of DEGs in primary melanoma vs normal skin",
    x = "Log Fold Change (logFC)",
    y = "Gene Symbol",
    color = "-Log10 Adjusted P-Value"
  ) +
  theme(legend.position = "top")
ggsave("analysis/2024-06-23_Primary_vs_normal.pdf", PT_DEG_plot)


```


### mel_dfci_2019 data cleaning
```{r}
MSK = read.table("data/mel_dfci_2019/data_mrna_seq_tpm.txt", sep = "\t", header = T)
# Create a named vector to map incorrect symbols to the correct ones
correct_mapping <- c(
  "1-Mar" = "MARCH1", "10-Mar" = "MARCH10", "11-Sep" = "SEPT11", 
  "1-Sep" = "SEPT1", "2-Mar" = "MARCH2", "2-Sep" = "SEPT2",
  "3-Mar" = "MARCH3", "3-Sep" = "SEPT3", "4-Sep" = "SEPT4",
  "5-Mar" = "MARCH5", "5-Sep" = "SEPT5", "6-Mar" = "MARCH6", 
  "6-Sep" = "SEPT6", "7-Mar" = "MARCH7", "7-Sep" = "SEPT7",
  "8-Mar" = "MARCH8", "8-Sep" = "SEPT8", "9-Mar" = "MARCH9", 
  "9-Sep" = "SEPT9", "10-Sep" = "SEPT10"
)

# Replace the incorrect symbols with the correct ones
MSK$Hugo_Symbol <- ifelse(MSK$Hugo_Symbol %in% names(correct_mapping), 
                          correct_mapping[MSK$Hugo_Symbol], 
                          MSK$Hugo_Symbol)

# Verify the correction
head(MSK$Hugo_Symbol, 25)
MSK$Entrez_Gene_Id = NULL
MSK = MSK %>% group_by(Hugo_Symbol) %>%
  summarize(across(everything(), sum)) %>% 
  filter(Hugo_Symbol > 0) %>% as.data.frame()


row.names(MSK) <- MSK$Hugo_Symbol
MSK$Hugo_Symbol = NULL

for (col in colnames(MSK)) {
  sum_fpkm <- sum(MSK[[col]])  # Use double brackets to get the column correctly
  MSK[[col]] <- (MSK[[col]] / sum_fpkm) * 1e6
}

GoI = c("EBI3", "KLK8", "MZB1", "NCKAP1L", "SAMSN1")

write.table(cbind(Gene = row.names(MSK), MSK), file = "data/2024-08-16_MSK_TPM.txt", sep = "\t", row.names = F, quote = F)


```


### mel_dfci_2019 analysis
```{r}

MSK_exp = read.table("data/2024-08-16_MSK_TPM.txt", header = T, sep = "\t", row.names = 1) %>% filter(row.names(.) %in% IRG)

MSK_exp_T = MSK_exp %>% (t) %>% as.data.frame()

MSK_exp_T = cbind(PatientID = row.names(MSK_exp_T), MSK_exp_T)

MSK_fraction = read.table("data/2024-08-16_MSK_GEP/CIBERSORTxGEP_Job48_Fractions-Adjusted.txt", header = T, sep = "\t", row.names = 1)

MSK_fraction = cbind(PatientID = row.names(MSK_fraction), MSK_fraction)

MSK_fraction_exp = merge(MSK_exp_T, MSK_fraction, by = "PatientID")

# Initialize a results data frame
results_MSK <- data.frame(Gene = character(), Correlation = numeric(), P_Value = numeric(), stringsAsFactors = FALSE)

# Loop through each gene column (excluding the T_CD8_fraction column)
for (gene in colnames(MSK_fraction_exp)[colnames(MSK_fraction_exp) %in% IRG]) {
  
  # Perform correlation test
  cor_test <- cor.test(log(MSK_fraction_exp[[gene]]+1), log(MSK_fraction_exp$T.cells.CD8 + 0.00001))
  
  # Store the results
  results_MSK <- rbind(results_MSK, data.frame(Gene = gene, Correlation = cor_test$estimate, P_Value = cor_test$p.value))
}

# Print the results
results_MSK$adj_p <- p.adjust(results_MSK$P_Value, method = "BH")
# Create a column for -log10(P-value)
results_MSK$logP_adj <- -log10(results_MSK$adj_p)

# Define your gene list to show
GeneShow <- c("B2M", GoI, "CD8A", "CD8B", "IL2RB", "PRF1", "JAK3", "PDCD1", "TNFSF10")

# Create the volcano plot
cor_plot = ggplot(results_MSK, aes(x = Correlation, y = logP_adj)) +
  geom_point(aes(color = logP_adj > -log10(0.05)), size = 1) + # Highlight significant points
  geom_label_repel(
    data = subset(results_MSK, logP_adj > -log10(0.05) & Gene %in% GeneShow), 
    aes(label = Gene), 
    size = 4, 
    #box.padding = 0.2,
    min.segment.length = 0,
    segment.color = 'grey50', # Line color
    segment.size = 0.7,  # Line thickness
    force = 10,           # Increase force to push labels away
    max.overlaps = Inf,   # Allow more overlaps before removing labels
    xlim = c(-Inf, Inf), ylim = c(-Inf, Inf)
  ) + 
  theme_minimal() +
  coord_cartesian(clip = "off") +
  labs(title = "Volcano Plot",
       x = "Correlation",
       y = "-log10(p.adj)",
       color = "Significant") +
  scale_color_manual(values = c("black", "red")) # Red for significant, black for non-significant


```


### Read scRNA-GSE115978
```{r}
# Read the annotations file
annotations <- read.table("data/2024-06-21_melanoma_scRNASeq/GSE115978/GSE115978_cell.annotations.csv.gz", header = T, row.names = 1, sep = ",")

annotations = annotations %>% mutate(CellType = case_when(
    cell.types %in% c("T.CD4", "T.CD8", "T.cell") ~ "T.cell",
    cell.types == "?" ~ "not defined",
    TRUE ~ cell.types))

# Read the counts file
counts <- read.table("data/2024-06-21_melanoma_scRNASeq/GSE115978/GSE115978_counts.csv.gz", header = T, row.names = 1, sep = ",")

# Create Seurat object from counts data
GSE115978_obj <- CreateSeuratObject(counts = counts, project = "GSE115978")

# Add cell annotations as metadata
GSE115978_obj <- AddMetaData(object = GSE115978_obj, metadata = annotations)


GSE115978_obj = subset(GSE115978_obj, nCount_RNA < 3e6 & CellType != "not defined")



# Normalizing the data
GSE115978_obj <- NormalizeData(GSE115978_obj)

# Identifying the 2000 most variable genes
GSE115978_obj <- FindVariableFeatures(GSE115978_obj, selection.method = "vst", nfeatures = 3000)

# Scaling the data
seurat_obj <- ScaleData(seurat_obj, features = rownames(seurat_obj))

# Running PCA
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(object = seurat_obj))

# Running UMAP
seurat_obj <- RunUMAP(seurat_obj, dims = 1:15)

seurat_obj_filter = subset(seurat_obj, nCount_RNA < 3e6 & cell.types != "?")

# Plotting UMAP
seurat_UMAP <- DimPlot(seurat_obj, reduction = "umap", group.by = "cell.types")

# Run UMAP
seurat_obj <- RunUMAP(seurat_obj, dims = 1:15)

# Visualize UMAP with cell identities
DimPlot(seurat_obj, reduction = "umap", group.by = "cell_type")  # Replace "ce


# Verify the Seurat object
print(seurat_obj)

# Save the Seurat object


# Set identities based on 'malignant.1.no.2.yes.0.unresolved.'
seurat_obj_filter <- SetIdent(seurat_obj_filter, value = seurat_obj_filter@meta.data$cell.types)

GOI = c("EBI3", "KLK8", "MZB1", "NCKAP1L", "SAMSN1")
FeaturePlot(seurat_obj_filter, features = GOI)

seurat_UMAP <- DimPlot(seurat_obj_filter, reduction = "umap", group.by = "cell.types")
saveRDS(seurat_obj, file = "GSE115978_seurat_obj.rds")

# Find markers between group 1 (non-malignant) and group 2 (malignant)
diff_genes <- FindMarkers(seurat_obj_filter, ident.1 = "Mal")

write.table(diff_genes, file = "data/2024-06-21_melanoma_scRNASeq/GSE115978/GSE115978_differential_genes.txt", sep = "\t", quote = FALSE, row.names = TRUE)
```


### Read scRNA-GSE72056
```{r}
# Step 1: Read the data

# Read in entire file
GSE72056 <- read.delim("data/2024-06-21_melanoma_scRNASeq/GSE72056/GSE72056_melanoma_single_cell_revised_v2.txt", header = T, stringsAsFactors = F)

# Pull only gene information
GSE72056_genes <- GSE72056[-1:-3,]

# Duplicate gene names so make names unique (up to you how you want to deal with this part)
gene_list <- GSE72056_genes %>% 
  pull("Cell") %>% 
  make.unique(sep = ".")

# Add back unique rownames
rownames(GSE72056_genes) <- gene_list

# Remove Column of gene names
GSE72056_genes <- GSE72056_genes[, -1]

# Pull meta data columns from original data
GSE72056_meta <- GSE72056[1:3,]

# Make rownames equal to column 1 values
rownames(GSE72056_meta) <- GSE72056_meta[, 1]

# Remove column 1
GSE72056_meta <- GSE72056_meta[, -1]

# Transpose meta data as Seurat expects meta data to have cell names as rows and meta data values as columns
GSE72056_meta_transpose <- data.frame(t(GSE72056_meta)) %>% mutate(CellType = case_when(
    malignant.1.no.2.yes.0.unresolved. == "0" ~ "not defined",
    malignant.1.no.2.yes.0.unresolved. == "2" ~ "Mal",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "1" ~ "T.cell",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "2" ~ "B.cell",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "3" ~ "Macrophage",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "4" ~ "Endo.",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "5" ~ "CAF",
    non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK. == "6" ~ "NK"))

GSE72056_meta_transpose$CellType[is.na(GSE72056_meta_transpose$CellType)] <- "not defined"
# Create Seurat Object
GSE72056_obj <- CreateSeuratObject(counts = GSE72056_genes, meta.data = GSE72056_meta_transpose)

GSE72056_obj = subset(GSE72056_obj, CellType != "not defined")

# Normalizing the data
GSE72056_obj <- NormalizeData(GSE72056_obj)

# Identifying the 2000 most variable genes
GSE72056_obj <- FindVariableFeatures(GSE72056_obj)

# Scaling the data
GSE72056_obj <- ScaleData(GSE72056_obj, features = rownames(GSE72056_seurat))

# Running PCA
GSE72056_seurat <- RunPCA(GSE72056_seurat, features = VariableFeatures(object = GSE72056_seurat))

# Running UMAP
GSE72056_seurat <- RunUMAP(GSE72056_seurat, dims = 1:15)

# Plotting UMAP
GSE72056_UMAP <- DimPlot(GSE72056_seurat, reduction = "umap", group.by = "non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK.")
GSE72056_UMAP2 <- DimPlot(GSE72056_seurat, reduction = "umap", group.by = "malignant.1.no.2.yes.0.unresolved.")
UMAPPlot

GSE72056_seurat = subset(GSE72056_seurat, malignant.1.no.2.yes.0.unresolved. != 0)
# Set identities based on 'malignant.1.no.2.yes.0.unresolved.'
GSE72056_seurat <- SetIdent(GSE72056_seurat, value = GSE72056_seurat@meta.data$malignant.1.no.2.yes.0.unresolved.)


# Find markers between group 1 (non-malignant) and group 2 (malignant)
diff_genes <- FindMarkers(GSE72056_seurat, ident.1 = 1, ident.2 = 2)
FeaturePlot(GSE72056_seurat, features = GOI)

write.table(diff_genes, file = "data/2024-06-21_melanoma_scRNASeq/GSE72056_differential_genes.txt", sep = "\t", quote = FALSE, row.names = TRUE)
GSE72056_UMAP2
```

### Perform Merge of the two scRNA data
```{r}
GSE72056_obj$Method = "GSE72056"
GSE115978_obj$Method = "GSE115978"
combined_obj <- merge(GSE72056_obj, y = GSE115978_obj)

# run standard analysis workflow
combined_obj <- NormalizeData(combined_obj)
combined_obj <- FindVariableFeatures(combined_obj)
combined_obj <- ScaleData(combined_obj)
combined_obj <- RunPCA(combined_obj)

#ElbowPlot(combined_obj)
combined_obj <- RunUMAP(combined_obj, dims = 1:15, reduction = "pca", reduction.name = "umap.unintegrated")
combined_obj_Dim_NoInt = DimPlot(combined_obj, reduction = "umap.unintegrated", group.by = c("Method", "CellType"))

# Integrate
combined_obj_int <- IntegrateLayers(object = combined_obj, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)
#combined_obj_int <- JoinLayers(combined_obj_int[["RNA"]])

combined_obj_int <- FindNeighbors(combined_obj_int, dims = 1:15, reduction = "integrated.cca")

combined_obj_int <- FindClusters(combined_obj_int, resolution = 0.2)
combined_obj_int <- RunUMAP(combined_obj_int, dims = 1:15, reduction = "integrated.cca")

combined_obj_int_method = DimPlot(combined_obj_int, reduction = "umap", group.by = c("Method"))
combined_obj_int_celltype = DimPlot(combined_obj_int, reduction = "umap", group.by = c("CellType"),  label = T)
combined_obj_int_celltype2 = DimPlot(combined_obj_int, reduction = "umap", group.by = c("cell.types"),  label = T)
combined_obj_int_celltype3 = DimPlot(combined_obj_int, reduction = "umap", group.by = c("non.malignant.cell.type..1.T.2.B.3.Macro.4.Endo..5.CAF.6.NK."),  label = T)

combined_obj_int_cluster = DimPlot(combined_obj_int, reduction = "umap", label = T)

combined_obj_int_annotation <- data.frame(cluster_ID = combined_obj_int@active.ident, row.names = rownames(combined_obj_int@meta.data)) %>% mutate(CellType_int = case_when(
  cluster_ID %in% c("0", "9") ~ "T_CD8", 
  cluster_ID == "1" ~ "T_CD4", 
  cluster_ID %in% c("2", "4", "6", "7", "10") ~ "Melanoma", 
  cluster_ID == "3" ~ "B_cell", 
  cluster_ID == "5" ~ "Macrophage",
  cluster_ID == "8" ~ "NK",
  cluster_ID == "11" ~ "CAF",
  cluster_ID %in% c("12","13") ~ "Endothelial"))
  
combined_obj_int = AddMetaData(combined_obj_int, metadata = combined_obj_int_annotation, col.name = "CellType_int")
combined_obj_int_celltype_final = DimPlot(combined_obj_int, reduction = "umap", group.by = c("CellType_int"), label = T) + NoLegend()
c("T_cell", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
    "NK", "DC", "Platelet")



MZB_Feature = FeaturePlot(combined_obj_int, features = GoI, reduction = "umap")

MZB_Dot = DotPlot(combined_obj_int, features = GoI,  dot.scale = 8, group.by = "CellType_int") + RotatedAxis()

saveRDS(combined_obj_int, "data/2024-09-09_melanoma_integrated.rds")

## Make more plots
combined_obj = readRDS("data/2024-09-09_melanoma_integrated.rds")
SERB4_Feature = FeaturePlot(combined_obj, features = "SERPINB4", reduction = "umap")
SERB4_Dot = DotPlot(combined_obj, features = "SERPINB4",  dot.scale = 8, group.by = "CellType_int") + RotatedAxis()

```

### Print fraction plot
```{r}


# Load the data from the Excel file
file_path <- "manuscript/TableS6_CellFractionofMelanomaTumor.xlsx"
cell_fraction_data <- read_excel(file_path, sheet = "TableS6_Cell_Fraction")

cell_fraction_data$Mixture <- gsub("Sample", "", cell_fraction_data$Mixture)


# Sort the data by 'Malignant' fraction in ascending order
cell_fraction_data <- cell_fraction_data %>%
  arrange(`Malignant`)

# Reshape the data into long format
cell_fraction_long <- melt(cell_fraction_data, id.vars = "Mixture", 
                           measure.vars = colnames(cell_fraction_data)[2:9], 
                           variable.name = "Cell_Type", value.name = "Cell_Fraction")

# Ensure the Mixture (sample IDs) are a factor and ordered by 'Malignant'
cell_fraction_long$Mixture <- factor(cell_fraction_long$Mixture, levels = cell_fraction_data$Mixture)

# Reorder Cell_Type so that 'Malignant' comes first, so it appears on top of the stack
cell_fraction_long$Cell_Type <- factor(cell_fraction_long$Cell_Type, 
                                       levels = c("Malignant", setdiff(unique(cell_fraction_long$Cell_Type), "Malignant")))

# Create the stacked bar plot
p = ggplot(cell_fraction_long, aes(x = Mixture, y = Cell_Fraction, fill = Cell_Type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Cell Fractions by Sample ID (Sorted by Malignant)", 
       x = "Sample ID", 
       y = "Cell Fraction (Stacked)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_brewer(palette = "Set3")
```

